---
title: "The end of COVID-19 is coming"
author: "Ilhak Park"
date: "21/06/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
supervisor: Bilgehan Erdem, Ph.D.
---
#### 0. Library
```{r}
library(rio)
library(forecast)
library(tseries)
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(zoo)
```


#### 1a). Pre-process Data - Loading Data
```{r}
dataset <- read.csv(file= "owid-covid-data.csv", stringsAsFactors =FALSE, header=TRUE)
```

#### 1b). Pre-process Data - Data Description
```{r}
dim(dataset)
head(dataset)
tail(dataset)
summary(dataset)
str(dataset)
sapply(dataset, class)
```

#### 1c). Pre-process Data - Selecting Variables & Observations
```{r}
#The class of Date is character, but the class should be date.
dataset$date <- as.Date(dataset$date)


#Creating a subset with only four columns that we need.
dataSubset <- data.frame("date" = dataset$date, "location" = dataset$location, "new_cases" = dataset$new_cases, "new_deaths" = dataset$new_deaths)
dataSubset 

#The period of the study is from March 1st, 2020 to June 1st, 2021.
China<-dataSubset %>% filter (location =='China', date>= '2020-03-01',date<='2021-06-01')
US<-dataSubset %>% filter (location =='United States', date>= '2020-03-01',date<='2021-06-01')
India<-dataSubset %>% filter (location =='India', date>= '2020-03-01',date<='2021-06-01')
UK<-dataSubset %>% filter (location =='United Kingdom', date>= '2020-03-01',date<='2021-06-01')
Germany<-dataSubset %>% filter (location =='Germany', date>= '2020-03-01',date<='2021-06-01')
Belgium<-dataSubset %>% filter (location =='Belgium', date>= '2020-03-01',date<='2021-06-01')
Netherlands<-dataSubset %>% filter (location =='Netherlands', date>= '2020-03-01',date<='2021-06-01')
```

#### 2a). EDA - Missing Values & Combine EU countries
```{r}
#There are missing values, in the new deaths variable, before first death cases occur, so we are replacing NA to 0.
India[is.na(India)] <-0
UK[is.na(UK)] <-0
Germany[is.na(Germany)] <-0
Belgium[is.na(Belgium)] <-0
Netherlands[is.na(Netherlands)] <-0

#Combine Germany Belgium and Netherlands as EU
EUCounts <- Germany[3:4] + Belgium[3:4] + Netherlands[3:4]
EUDates <- Germany[1]
EULocation<- matrix(rep('EU',458),458,1)
colnames(EULocation) <-c('location')
EU <- cbind(EUDates,EULocation,EUCounts)
```

#### 2b). EDA - Correlation between new cases and new deaths 
```{r}
cor(US$new_cases,US$new_deaths)
cor(China$new_cases,China$new_deaths)
cor(India$new_cases, India$new_deaths)
cor(UK$new_cases, UK$new_deaths)
cor(EU$new_cases, EU$new_deaths)
#New cases and new deaths have a positive correlation in all countries. Since new cases reach zero, new deaths will also be zero. In this study we will focus on forecasting new cases only. 
```

#### 2c). EDA - Negative values (false postive cases)
```{r}
#Negative values are observed in new cases for China and UK. The negative values are false positives due to negative confirmatory PCR.
summary(US)
summary(China)
summary(India)
summary(UK)
summary(EU)

#China has a false positive case, so we will change the new case value on 2020-06-03 from -1 to 0, and reduce the new case value on 2020-06-02 from 1 to 0.
China[95,3] <- 0
China[94,3] <- 0
summary(China)

#UK has thousands of false positive cases, from the beginning, and updated the data by reducing 8,010 cases on April 4, 2021.
#In order to have an accurate data, we will import the data released by UK government. (https://coronavirus.data.gov.uk/details/cases)
UK_data <- read.csv(file= "UK_data.csv", stringsAsFactors =FALSE, header=TRUE)
Subset_UK <- data.frame("date" = UK_data$date, "location" = UK_data$areaName, "new_cases" = UK_data$newCasesBySpecimenDate)
Subset_UK <- Subset_UK[order(Subset_UK$date),]
UK<-Subset_UK %>% filter (date>= '2020-03-01',date<='2021-06-01')
summary(UK)
```


#### 2d). EDA - Decomposition check
```{r}
#It is difficult to tell whether the data has a weekly seasonality or a monthly seasonality.
#Creating 2 different models to compare

#Models with frequency = 7
ts_US <- ts(US$new_cases, frequency = 7)
ts_US
decompose_US <- decompose(ts_US, "additive")
plot(decompose_US)

ts_China <- ts(China$new_cases, frequency = 7)
decompose_China <- decompose(ts_China, "additive")
plot(decompose_China)

ts_India <- ts(India$new_cases, frequency = 7)
decompose_India <- decompose(ts_India, "additive")
plot(decompose_India)

ts_UK <- ts(UK$new_cases, frequency = 7)
decompose_UK <- decompose(ts_UK, "additive")
plot(decompose_UK)

ts_EU <- ts(EU$new_cases, frequency = 7)
decompose_EU <- decompose(ts_EU, "additive")
plot(decompose_EU)

#Models with freqeuncy =12
ts_US12 <- ts(US$new_cases, frequency = 12)
ts_US12
decompose_US12 <- decompose(ts_US12, "additive")
plot(decompose_US12)

ts_China12 <- ts(China$new_cases, frequency = 12)
decompose_China12 <- decompose(ts_China12, "additive")
plot(decompose_China12)

ts_India12 <- ts(India$new_cases, frequency = 12)
decompose_India12 <- decompose(ts_India12, "additive")
plot(decompose_India12)

ts_UK12 <- ts(UK$new_cases, frequency = 12)
decompose_UK12 <- decompose(ts_UK12, "additive")
plot(decompose_UK12)

ts_EU12 <- ts(EU$new_cases, frequency = 12)
decompose_EU12 <- decompose(ts_EU12, "additive")
plot(decompose_EU12)
```
#### 2e). EDA - Stationary check
```{r}
acf(ts_US)
#ACF is decaying slowly and remains above the significance range (blue lines), which means it is non-stationary
adf.test(as.matrix(ts_US))
#Conducting Augmented Dickey-Fuller Test to check if the data is stationary
#p-value is greater than 0.05, which means it is non-stationary again.
acf(ts_China, lag.max = 34)
adf.test(as.matrix(ts_China))
#ts_China is stationary
acf(ts_India, lag.max = 34)
adf.test(as.matrix(ts_India))
#ts_India seems non-stationary from its ACF, but passed the ADF test
acf(ts_UK)
adf.test(as.matrix(ts_UK))
#ts_UK is non-stationary
acf(ts_EU)
adf.test(as.matrix(ts_EU))
#ts_EU is non-stationary


#Same stationary check for models with frequency =12
acf(ts_US12)
adf.test(as.matrix(ts_US12))
#ts_US12 is non-stationary
acf(ts_China12, lag.max = 34)
adf.test(as.matrix(ts_China12))
#ts_China12 is stationary
acf(ts_India12, lag.max = 34)
adf.test(as.matrix(ts_India12))
#ts_India12 seems non-stationary from its ACF, but passed the ADF test
acf(ts_UK12)
adf.test(as.matrix(ts_UK12))
#ts_UK12 is non-stationary
acf(ts_EU12)
adf.test(as.matrix(ts_EU12))
#ts_EU is non-stationary
```

#### 3a-1). Modelling - US (training set of 2020/03 - 2021/06)
```{r}
#finding the best ARIMA model with respect to AIC value
model_US <- auto.arima(ts_US, ic="aic", trace =TRUE, approximation = F)
model_US
#checking stationary after seasonal differencing
acf(ts(model_US$residuals))
pacf(ts(model_US$residuals))
#ACF shows exponential decays and indicates it is stationary

#forecasting 1 month ahead (4-weekly forecast) with a 95% interval
forecast_US<-forecast(model_US, h=4, level=c(95))
forecast_US
plot(forecast_US)
summary(forecast_US)
accuracy(forecast_US)


#actual new cases in US on July 1, 2021
(dataset %>% filter (location =='United States', date=='2021-07-01'))[6]
#14463
#difference between the actual and predicted
abs(14463 - 8893.088)
#5569
```

#### 3a-2). Modeling - US (training set of 2021/01 - 2021/06)
```{r}
US_2021 <- US %>% filter (location =='United States', date>= '2021-01-01',date<='2021-06-01')

#checking decomposition and stationary
ts_US_2021 <- ts(US_2021$new_cases,frequency =7)
ts_US_2021
decompose_US_2021 <- decompose(ts_US_2021, "additive")
plot(decompose_US_2021)
acf(ts_US_2021)
pacf(ts_US_2021)
model_US_2021 <- auto.arima(ts_US_2021, ic="aic", trace =TRUE, approximation = F)
model_US_2021
acf(ts(model_US_2021$residuals))
pacf(ts(model_US_2021$residuals))

#forecasting 1 month ahead (4-weekly forecast) with a 95% interval
forecast_US_2021<-forecast(model_US_2021, h=4, level=c(95))
forecast_US_2021
plot(forecast_US_2021)
summary(forecast_US_2021)
accuracy(forecast_US_2021)

#difference between the actual and predicted
abs(14463 - 4236.653)
#10226.35
```

#### 3a-3). Modelling - US (training set of 2020/03 - 2021/06)
```{r}
#finding the best ARIMA model with respect to AIC value
model_US12 <- auto.arima(ts_US12, ic="aic", trace =TRUE, approximation = F)
model_US12
#checking stationary after seasonal differencing
acf(ts(model_US12$residuals))
pacf(ts(model_US12$residuals))
#ACF shows exponential decays and indicates it is stationary

#forecasting 1 month ahead with a 95% interval
forecast_US12<-forecast(model_US12, h=1, level=c(95))
forecast_US12
plot(forecast_US12)



#actual new cases in US on July 1, 2021
(dataset %>% filter (location =='United States', date=='2021-07-01'))[6]
#14463
#difference between the actual and predicted
abs(14463 - 19618.77)
```

#### 3a-4). Modeling - US (training set of 2021/01 - 2021/06)
```{r}
US12_2021 <- US %>% filter (location =='United States', date>= '2021-01-01',date<='2021-06-01')

#checking decomposition and stationary
ts_US12_2021 <- ts(US12_2021$new_cases,frequency =12)
ts_US12_2021
decompose_US12_2021 <- decompose(ts_US12_2021, "additive")
plot(decompose_US12_2021)
acf(ts_US12_2021)
pacf(ts_US12_2021)
model_US12_2021 <- auto.arima(ts_US12_2021, ic="aic", trace =TRUE, approximation = F)
model_US12_2021
acf(ts(model_US12_2021$residuals))
pacf(ts(model_US12_2021$residuals))

#forecasting 1 month ahead with a 95% interval
forecast_US12_2021<-forecast(model_US12_2021, h=1, level=c(95))
forecast_US12_2021
plot(forecast_US12_2021)
summary(forecast_US12_2021)

#difference between the actual and predicted
abs(14463 - 19318.81)
```

#### 3b-1). Modeling - China (training set of 2020/03 - 2021/06)
```{r}
#finding the best ARIMA model with respect to AIC value
model_China <- auto.arima(ts_China, ic="aic", trace =TRUE, approximation = F)
model_China
#checking stationary after seasonal differencing
acf(ts(model_China$residuals))
pacf(ts(model_China$residuals))
#ACF shows exponential decays and indicates it is stationary

#forecasting 1 month ahead with a 95% interval
forecast_China<-forecast(model_China, h=4, level=c(95))
forecast_China
plot(forecast_China)
summary(forecast_China)
accuracy(forecast_China)


#actual new cases in China on July 1, 2021
(dataset %>% filter (location =='China', date=='2021-08-01'))[6]
#18
#difference between the actual and predicted
abs(18 - 29.76)
```

#### 3b-2). Modeling - China (training set of 2021/01 - 2021/06)
```{r}
China_2021 <- China %>% filter ( date>= '2021-01-01',date<='2021-06-01')

#checking decomposition and stationary
ts_China_2021 <- ts(China_2021$new_cases,frequency =7)
ts_China_2021
decompose_China_2021 <- decompose(ts_China_2021, "additive")
plot(decompose_China_2021)
acf(ts_China_2021)
pacf(ts_China_2021)
model_China_2021 <- auto.arima(ts_China_2021, ic="aic", trace =TRUE, approximation = F)
model_China_2021
acf(ts(model_China_2021$residuals))
pacf(ts(model_China_2021$residuals))

#forecasting 1 month ahead (4-weekly forecast) with a 95% interval
forecast_China_2021<-forecast(model_China_2021, h=4, level=c(95))
forecast_China_2021
plot(forecast_China_2021)
summary(forecast_China_2021)
accuracy(forecast_China_2021)

#actual new cases in China on July 1, 2021
(dataset %>% filter (location =='China', date=='2021-07-01'))[6]
#18
abs(18 - 29.7)
```

#### 3b-3). Modeling - China (training set of 2020/03 - 2021/06)
```{r}
#finding the best ARIMA model with respect to AIC value
model_China12 <- auto.arima(ts_China12, ic="aic", trace =TRUE, approximation = F)
model_China12
#checking stationary after seasonal differencing
acf(ts(model_China12$residuals))
pacf(ts(model_China12$residuals))
#ACF shows exponential decays and indicates it is stationary

#forecasting 1 month ahead with a 95% interval
forecast_China12<-forecast(model_China12, h=1, level=c(95))
forecast_China12
plot(forecast_China12)
summary(forecast_China12)

#actual new cases in China on July 1, 2021
(dataset %>% filter (location =='China', date=='2021-07-01'))[6]
#18
#difference between the actual and predicted
abs(18 - 23.33)
```

#### 3b-4). Modeling - China (training set of 2021/01 - 2021/06)
```{r}
China12_2021 <- China %>% filter ( date>= '2021-01-01',date<='2021-06-01')

#checking decomposition and stationary
ts_China12_2021 <- ts(China12_2021$new_cases,frequency =12)
ts_China12_2021
decompose_China12_2021 <- decompose(ts_China12_2021, "additive")
plot(decompose_China12_2021)
acf(ts_China12_2021)
pacf(ts_China12_2021)
model_China12_2021 <- auto.arima(ts_China12_2021, ic="aic", trace =TRUE, approximation = F)
model_China12_2021
acf(ts(model_China12_2021$residuals))
pacf(ts(model_China12_2021$residuals))

#forecasting 1 month ahead with a 95% interval
forecast_China12_2021<-forecast(model_China12_2021, h=1, level=c(95))
forecast_China12_2021
plot(forecast_China12_2021)
summary(forecast_China12_2021)

#actual new cases in China on July 1, 2021
(dataset %>% filter (location =='China', date=='2021-07-01'))[6]
#18
abs(18 - 19.86)
```

#### 3c-1). Modeling - India (training set of 2020/03 - 2021/06)
```{r}
#finding the best ARIMA model with respect to AIC value
model_India <- auto.arima(ts_India, ic="aic", trace =TRUE, approximation = FALSE)
model_India
#checking stationary after seasonal differencing
acf(ts(model_India$residuals))
pacf(ts(model_India$residuals))
#ACF shows exponential decays and indicates it is stationary

#forecasting 1 month ahead with a 95% interval
forecast_India<-forecast(model_India, h=4, level=c(95))
forecast_India
plot(forecast_India)
summary(forecast_India)
accuracy(forecast_India)


#actual new cases in India on July 1, 2021
(dataset %>% filter (location =='India', date=='2021-07-01'))[6]
#46617
#difference between the actual and predicted
abs(46617 -98886.52)
```

#### 3c-2). Modeling - India (training set of 2021/01 - 2021/06)
```{r}
India_2021 <- India %>% filter ( date>= '2021-01-01',date<='2021-06-01')

#checking decomposition and stationary
ts_India_2021 <- ts(India_2021$new_cases,frequency =7)
ts_India_2021
decompose_India_2021 <- decompose(ts_India_2021, "additive")
plot(decompose_India_2021)
acf(ts_India_2021)
pacf(ts_India_2021)
model_India_2021 <- auto.arima(ts_India_2021, ic="aic", trace =TRUE, approximation = F)
model_India_2021
acf(ts(model_India_2021$residuals))
pacf(ts(model_India_2021$residuals))

#forecasting 1 month ahead (4-weekly forecast) with a 95% interval
forecast_India_2021<-forecast(model_India_2021, h=4, level=c(95))
forecast_India_2021
plot(forecast_India_2021)

#actual new cases in India on July 1, 2021
(dataset %>% filter (location =='India', date=='2021-07-01'))[6]
#46617
#difference between the actual and predicted
abs(46617 - 91799.52)
```

#### 3c-3). Modeling - India (training set of 2020/03 - 2021/06)
```{r}
#finding the best ARIMA model with respect to AIC value
model_India12 <- auto.arima(ts_India12, ic="aic", trace =TRUE, approximation = FALSE)
model_India12
#checking stationary after seasonal differencing
acf(ts(model_India12$residuals))
pacf(ts(model_India12$residuals))
#ACF shows exponential decays and indicates it is stationary

#forecasting 1 month ahead with a 95% interval
forecast_India12<-forecast(model_India12, h=1, level=c(95))
forecast_India12
plot(forecast_India12)
summary(forecast_India12)


#actual new cases in India on July 1, 2021
(dataset %>% filter (location =='India', date=='2021-08-01'))[6]
#46617
#difference between the actual and predicted
abs(46617 -143780.20)
```

#### 3c-4). Modeling - India (training set of 2021/01 - 2021/06)
```{r}
India12_2021 <- India %>% filter ( date>= '2021-01-01',date<='2021-06-01')

#checking decomposition and stationary
ts_India12_2021 <- ts(India12_2021$new_cases,frequency =12)
ts_India12_2021
decompose_India12_2021 <- decompose(ts_India12_2021, "additive")
plot(decompose_India12_2021)
acf(ts_India12_2021)
pacf(ts_India12_2021)
model_India12_2021 <- auto.arima(ts_India12_2021, ic="aic", trace =TRUE, approximation = F)
model_India12_2021
acf(ts(model_India12_2021$residuals))
pacf(ts(model_India12_2021$residuals))

#forecasting 1 month ahead with a 95% interval
forecast_India12_2021<-forecast(model_India12_2021, h=1, level=c(95))
forecast_India12_2021
plot(forecast_India12_2021)
summary(forecast_India12_2021)

#actual new cases in India on July 1, 2021
(dataset %>% filter (location =='India', date=='2021-07-01'))[6]
#46617
#difference between the actual and predicted
abs(46617 - 118705)
```

#### 3d-1). Modeling - UK (training set of 2020/03 - 2021/06)
```{r}
#finding the best ARIMA model with respect to AIC value
model_UK <- auto.arima(ts_UK, ic="aic", trace =TRUE, approximation = F)
model_UK
#checking stationary after seasonal differencing
acf(ts(model_UK$residuals))
pacf(ts(model_UK$residuals))
#ACF shows exponential decays and indicates it is stationary

#forecasting 1 month ahead with a 95% interval
forecast_UK<-forecast(model_UK, h=4, level=c(95))
forecast_UK
plot(forecast_UK)
summary(forecast_UK)
accuracy(forecast_UK)

#actual new cases in UK on July 1, 2021 is 28071
#difference between the actual and predicted
abs(28071 - 3944.238)
```

#### 3d-2). Modeling - UK (training set of 2021/01 - 2021/06)
```{r}
UK_2021 <- UK %>% filter (date>= '2021-01-01',date<='2021-06-01')

#checking decomposition and stationary
ts_UK_2021 <- ts(UK_2021$new_cases,frequency =7)
ts_UK_2021
decompose_UK_2021 <- decompose(ts_UK_2021, "additive")
plot(decompose_UK_2021)
acf(ts_UK_2021)
pacf(ts_UK_2021)
model_UK_2021 <- auto.arima(ts_UK_2021, ic="aic", trace =TRUE, approximation = F)
model_UK_2021
acf(ts(model_UK_2021$residuals))
pacf(ts(model_UK_2021$residuals))

#forecasting 1 month ahead (4-weekly forecast) with a 95% interval
forecast_UK_2021<-forecast(model_UK_2021, h=4, level=c(95))
forecast_UK_2021
plot(forecast_UK_2021)
summary(forecast_UK_2021)
accuracy(forecast_UK_2021)

#actual new cases in UK on July 1, 2021 is 28071
#difference between the actual and predicted
abs(28071 - 4175.98)
```

#### 3d-3). Modeling - UK (training set of 2020/03 - 2021/06)
```{r}
#finding the best ARIMA model with respect to AIC value
model_UK12 <- auto.arima(ts_UK12, ic="aic", trace =TRUE, approximation = F)
model_UK12
#checking stationary after seasonal differencing
acf(ts(model_UK12$residuals))
pacf(ts(model_UK12$residuals))
#ACF shows exponential decays and indicates it is stationary

#forecasting 1 month ahead with a 95% interval
forecast_UK12<-forecast(model_UK12, h=1, level=c(95))
forecast_UK12
plot(forecast_UK12)
summary(forecast_UK12)

#actual new cases in UK on July 1, 2021 is 28071
#difference between the actual and predicted
abs(28071 - 4599.55)
```

#### 3d-4). Modeling - UK (training set of 2021/01 - 2021/06)
```{r}
UK12_2021 <- UK %>% filter (date>= '2021-01-01',date<='2021-06-01')

#checking decomposition and stationary
ts_UK12_2021 <- ts(UK12_2021$new_cases,frequency =12)
ts_UK12_2021
decompose_UK12_2021 <- decompose(ts_UK12_2021, "additive")
plot(decompose_UK12_2021)
acf(ts_UK12_2021)
pacf(ts_UK12_2021)
model_UK12_2021 <- auto.arima(ts_UK12_2021, ic="aic", trace =TRUE, approximation = F)
model_UK12_2021
acf(ts(model_UK12_2021$residuals))
pacf(ts(model_UK12_2021$residuals))

#forecasting 1 month ahead (4-weekly forecast) with a 95% interval
forecast_UK12_2021<-forecast(model_UK12_2021, h=1, level=c(95))
forecast_UK12_2021
plot(forecast_UK12_2021)
summary(forecast_UK12_2021)

#actual new cases in UK on July 1, 2021 is 28071
#difference between the actual and predicted
abs(28071 - 4283.79)
```

#### 3e-1). Modeling - EU (training set of 2020/03 - 2021/06)
```{r}
#finding the best ARIMA model with respect to AIC value
model_EU <- auto.arima(ts_EU, ic="aic", trace =TRUE, approximation = F)
model_EU
#checking stationary after seasonal differencing
acf(ts(model_EU$residuals))
pacf(ts(model_EU$residuals))
#ACF shows exponential decays and indicates it is stationary

#forecasting 1 month ahead with a 95% interval
forecast_EU<-forecast(model_EU, h=4, level=c(95))
forecast_EU
plot(forecast_EU)
summary(forecast_EU)
accuracy(forecast_EU)


#actual new cases in EU on July 1, 2021 is 2337
(dataset %>% filter (location =='Belgium', date=='2021-08-01'))[6] +(dataset %>% filter (location =='Netherlands', date=='2021-08-01'))[6] +(dataset %>% filter (location =='Germany', date=='2021-08-01'))[6]
#difference between the actual and predicted
abs(2337 - 7650.1)
```

#### 3e-2). Modeling - EU (training set of 2021/01 - 2021/06)
```{r}
EU_2021 <- EU %>% filter (date>= '2021-01-01',date<='2021-06-01')

#checking decomposition and stationary
ts_EU_2021 <- ts(EU_2021$new_cases,frequency =7)
ts_EU_2021
decompose_EU_2021 <- decompose(ts_EU_2021, "additive")
plot(decompose_EU_2021)
acf(ts_EU_2021)
pacf(ts_EU_2021)
model_EU_2021<- auto.arima(ts_EU_2021, ic="aic", trace =TRUE, approximation = F)
model_EU_2021
acf(ts(model_EU_2021$residuals))
pacf(ts(model_EU_2021$residuals))

#forecasting 1 month ahead (4-weekly forecast) with a 95% interval
forecast_EU_2021<-forecast(model_EU_2021, h=4, level=c(95))
forecast_EU_2021
plot(forecast_EU_2021)
summary(forecast_EU_2021)
accuracy(forecast_EU_2021)

#actual new cases in EU on July 1, 2021 is 2337
(dataset %>% filter (location =='Belgium', date=='2021-07-01'))[6] +(dataset %>% filter (location =='Netherlands', date=='2021-07-01'))[6] +(dataset %>% filter (location =='Germany', date=='2021-07-01'))[6]
#difference between the actual and predicted
abs(2337 - 5704.728)
```

#### 3e-3). Modeling - EU (training set of 2020/03 - 2021/06)
```{r}
#finding the best ARIMA model with respect to AIC value
model_EU12 <- auto.arima(ts_EU12, ic="aic", trace =TRUE, approximation = F)
model_EU12
#checking stationary after seasonal differencing
acf(ts(model_EU12$residuals))
pacf(ts(model_EU12$residuals))
#ACF shows exponential decays and indicates it is stationary

#forecasting 1 month ahead with a 95% interval
forecast_EU12<-forecast(model_EU12, h=1, level=c(95))
forecast_EU12
plot(forecast_EU12)
summary(forecast_EU12)

#actual new cases in EU on July 1, 2021 is 2337
(dataset %>% filter (location =='Belgium', date=='2021-07-01'))[6] +(dataset %>% filter (location =='Netherlands', date=='2021-07-01'))[6] +(dataset %>% filter (location =='Germany', date=='2021-07-01'))[6]
#difference between the actual and predicted
abs(2337 - 10339.34)
```

#### 3e-4). Modeling - EU (training set of 2021/01 - 2021/06)
```{r}
EU12_2021 <- EU %>% filter (date>= '2021-01-01',date<='2021-06-01')

#checking decomposition and stationary
ts_EU12_2021 <- ts(EU12_2021$new_cases,frequency =12)
ts_EU12_2021
decompose_EU12_2021 <- decompose(ts_EU12_2021, "additive")
plot(decompose_EU12_2021)
acf(ts_EU12_2021)
pacf(ts_EU12_2021)
model_EU12_2021<- auto.arima(ts_EU12_2021, ic="aic", trace =TRUE, approximation = F)
model_EU12_2021
acf(ts(model_EU12_2021$residuals))
pacf(ts(model_EU12_2021$residuals))

#forecasting 1 month ahead with a 95% interval
forecast_EU12_2021<-forecast(model_EU12_2021, h=1, level=c(95))
forecast_EU12_2021
plot(forecast_EU12_2021)
summary(forecast_EU12_2021)

#actual new cases in EU on July 1, 2021 is 2337
(dataset %>% filter (location =='Belgium', date=='2021-07-01'))[6] +(dataset %>% filter (location =='Netherlands', date=='2021-07-01'))[6] +(dataset %>% filter (location =='Germany', date=='2021-07-01'))[6]
#difference between the actual and predicted
abs(2337 - 12023.28)
```


#### 4a). US - Uncertainty Comparison
```{r}
#test set is 1 month of values after June 1st, 2021
US_testset<-dataSubset %>% filter (location =='United States', date>= '2021-06-02',date<='2021-07-01')

ts_US_testset <- ts(US_testset$new_cases, frequency =7)
ts_US_testset

#picking actual values of four 7-day
i<-1
actual_US<-ts_US_testset[1:(i+6)==(i+6)]
actual_US

accuracy(forecast_US, actual_US)
#mean absolute percentage error (MAPE) = 42.17
#root mean squared error (RMSE) = 5263.455

accuracy(forecast_US_2021, actual_US)
#mean absolute percentage error (MAPE) = 42.90
#root mean squared error (RMSE) = 5321.225

#picking an actual values after 30 days
i<-1
actual_US12<-ts_US_testset[30]
actual_US12

accuracy(forecast_US12, actual_US12)
#mean absolute percentage error (MAPE) = 35.65
#root mean squared error (RMSE) = 5155.773

accuracy(forecast_US12_2021, actual_US12)
#mean absolute percentage error (MAPE) = 33.57
#root mean squared error (RMSE) = 4855.801
```

#### 4b). China - Uncertainty Comparison
```{r}
#test set is 1 month of values after June 1st, 2021
China_testset<-dataSubset %>% filter (location =='China', date>= '2021-06-02',date<='2021-07-01')

ts_China_testset <- ts(China_testset$new_cases, frequency =7)
ts_China_testset

#picking actual values of four 7-day
i<-1
actual_China<-ts_China_testset[1:(i+6)==(i+6)]
actual_China

accuracy(forecast_China, actual_China)
#mean absolute percentage error (MAPE) = 75.13
#root mean squared error (RMSE) = 10.94

accuracy(forecast_China_2021, actual_China)
#mean absolute percentage error (MAPE) = 69.78
#root mean squared error (RMSE) = 10.85

#picking an actual value after 30 days
i<-1
actual_China12<-ts_China_testset[30]
actual_China12

accuracy(forecast_China12, actual_China12)
#mean absolute percentage error (MAPE) = 29.59
#root mean squared error (RMSE) = 5.3263

accuracy(forecast_China12_2021, actual_China12)
#mean absolute percentage error (MAPE) = 10.35
#root mean squared error (RMSE) = 1.86
```

#### 4c). India - Uncertainty Comparison
```{r}
#test set is 1 month of values after June 1st, 2021
India_testset<-dataSubset %>% filter (location =='India', date>= '2021-06-02',date<='2021-07-01')

ts_India_testset <- ts(India_testset$new_cases, frequency =7)
ts_India_testset

#picking actual values of four 7-day
i<-1
actual_India<-ts_India_testset[1:(i+6)==(i+6)]
actual_India

accuracy(forecast_India, actual_India)
#mean absolute percentage error (MAPE) = 91.39
#root mean squared error (RMSE) = 52880.40

accuracy(forecast_India_2021, actual_India)
#mean absolute percentage error (MAPE) = 82.69
#root mean squared error (RMSE) = 48218.16

#picking an actual value after 30 days
i<-1
actual_India12<-ts_India_testset[30]
actual_India12

accuracy(forecast_India12, actual_India12)
#mean absolute percentage error (MAPE) = 208.43
#root mean squared error (RMSE) = 97163.18

accuracy(forecast_India12_2021, actual_India12)
#mean absolute percentage error (MAPE) = 154.64
#root mean squared error (RMSE) = 72088.031
```

#### 4d). UK - Uncertainty Comparison
```{r}
#test set is 1 month of values after June 1st, 2021
UK_testset<-Subset_UK %>% filter ( date>= '2021-06-02',date<='2021-07-01')

ts_UK_testset <- ts(UK_testset$new_cases, frequency =7)
ts_UK_testset

#picking actual values of four 7-day
i<-1
actual_UK<-ts_UK_testset[1:(i+6)==(i+6)]
actual_UK

accuracy(forecast_UK, actual_UK)
#mean absolute percentage error (MAPE) = 63.49
#root mean squared error (RMSE) = 13806.76

accuracy(forecast_UK_2021, actual_UK)
#mean absolute percentage error (MAPE) = 61.48
#root mean squared error (RMSE) = 13663.83

#picking an actual value after 30 days
i<-1
actual_UK12<-ts_UK_testset[30]
actual_UK12

accuracy(forecast_UK12, actual_UK12)
#mean absolute percentage error (MAPE) = 83.61
#root mean squared error (RMSE) = 23471.451

accuracy(forecast_UK12_2021, actual_UK12)
#mean absolute percentage error (MAPE) = 84.74
#root mean squared error (RMSE) = 23787.206
```

#### 4e). EU - Uncertainty Comparison
```{r}
#test set is 1 month of values after June 1st, 2021
#Combine Germany Belgium and Netherlands as EU
Germany_testset<-dataSubset %>% filter (location =='Germany', date>= '2021-06-02',date<='2021-07-01')
Belgium_testset<-dataSubset %>% filter (location =='Belgium', date>= '2021-06-02',date<='2021-07-01')
Netherlands_testset<-dataSubset %>% filter (location =='Netherlands', date>= '2021-06-02',date<='2021-07-01')
EU_testset <- Germany_testset$new_cases +Belgium_testset$new_cases + Netherlands_testset$new_cases
EU_testset

ts_EU_testset <- ts(EU_testset, frequency =7)
ts_EU_testset

#picking actual values of four 7-day
i<-1
actual_EU<-ts_EU_testset[1:(i+6)==(i+6)]
actual_EU

accuracy(forecast_EU, actual_EU)
#mean absolute percentage error (MAPE) = 293.51
#root mean squared error (RMSE) = 7342.32

accuracy(forecast_EU_2021, actual_EU)
#mean absolute percentage error (MAPE) = 254.48
#root mean squared error (RMSE) = 6719.075

#picking an actual value after 30 days
i<-1
actual_EU12<-ts_EU_testset[30]
actual_EU12

accuracy(forecast_EU12, actual_EU12)
#mean absolute percentage error (MAPE) = 342.42
#root mean squared error (RMSE) = 8002.341

accuracy(forecast_EU12_2021, actual_EU12)
#mean absolute percentage error (MAPE) = 414.47
#root mean squared error (RMSE) = 9686.28
```


#### 5a). Result - US
```{r}
forecast_US12_2021<-forecast(model_US12_2021, h=1*12, level=c(95))
forecast_US12_2021
plot(forecast_US12_2021, ylim=c(0,300000), xlab="Number of months since January 2021", ylab="New COVID-19 cases", main="US - 1 Year Forecast using ARIMA")
```

#### 5b). Result - China
```{r}
forecast_China12_2021<-forecast(model_China12_2021, h=1*12, level=c(95))
forecast_China12_2021
plot(forecast_China12_2021, ylim=c(0,160), xlab="Number of months since January 2021", ylab="New COVID-19 cases", main="China - 1 Year Forecast using ARIMA")
```

#### 5c). Result - India
```{r}
forecast_India_2021<-forecast(model_India_2021, h=4*3, level=c(95))
forecast_India_2021
plot(forecast_India_2021,ylim=c(0,420000), xlab="Number of weeks since January 2021", ylab="New COVID-19 cases", main="India - 3 Months Forecast using ARIMA")
```
#### 3d-2). Modeling - UK (training set of 2021/01 - 2021/06)
```{r}
forecast_UK_2021<-forecast(model_UK_2021, h=4*7, level=c(95))
forecast_UK_2021
plot(forecast_UK_2021,ylim=c(0,80000), xlab="Number of weeks since January 2021", ylab="New COVID-19 cases", main="UK - 7 months forecast using ARIMA")
```

#### 5e). Result - EU
```{r}
forecast_EU_2021<-forecast(model_EU_2021, h=4*3, level=c(95))
forecast_EU_2021
plot(forecast_EU_2021,ylim=c(0,60000), xlab="Number of weeks since January 2021", ylab="New COVID-19 cases", main="EU - 3 months forecast using ARIMA")
```

